### 2016-11-22 practical_2/report/report_lastname.tex
update

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage[final]{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{fancyvrb,xcolor}
\usepackage{graphicx}

\title{Report for the Deep Learning Course Assignment 2 }

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Georgios Methenitis \\
  \texttt{georgios.methenitis@cwi.nl}
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
In this assignment I got familiar with TensorFlow. I created a single hidden layer network with ReLU activation function to learn features for the CIFAR10 dataset.
\end{abstract}

\section{Task 1}
\begin{enumerate}
\item 
\item There are two ways to declare variable in TensorFlow:
\begin{itemize}
\item \begin{small}
\begin{verbatim}
var = tf.get_variable("var_name", shape=[..., ...])
\end{verbatim}
\end{small}
\item \begin{small}
\begin{verbatim}
var = tf.Variable(tf.zeros([..., ...]), name="var_name")
\end{verbatim}
\end{small}
\end{itemize}

\item The command \texttt{tf.shape(x)} returns the shape of the input of tensor \texttt{x} in an 1-D vector, while \texttt{x.get\_shape()} return the size of the tensor.
\item
\item
\item
\item
\item
\item
\item
\end{enumerate}

\section{Task 2}
In this task I implemented the following functions in \texttt{mlp.py}, more specifically the \texttt{inference}, \texttt{loss}, and \texttt{accuracy}.

\begin{small}
\begin{verbatim}
  def inference(self, x):
    with tf.variable_scope('hidden', reuse=None):
      W = tf.get_variable("weights", shape=[x.get_shape()[1], self.n_hidden[0]],
       initializer=self.weight_initializer, regularizer=self.weight_regularizer)
      tf.histogram_summary("hidden_weights", W)
      b = tf.Variable(tf.zeros([self.n_hidden[0]]), name="bias")
      tf.histogram_summary("hidden_bias", b)

    with tf.variable_scope('output', reuse=None):
      w_out = tf.get_variable("weights", shape=[self.n_hidden[0], self.n_classes],
       initializer=self.weight_initializer, regularizer=self.weight_regularizer)
      tf.histogram_summary("output_weights", w_out)
      b_out = tf.Variable(tf.zeros([self.n_classes]), name="bias")
      tf.histogram_summary("output_bias", b_out)

    with tf.name_scope('relu_layer'):
      input = self.activation_fn(tf.matmul(x, W) + b)
    with tf.name_scope('linear_layer'):
      output = tf.matmul(input, w_out) + b_out

    return output

  def loss(self, logits, labels):

    loss = tf.reduce_mean(tf.nn.softmax_cross_entropy_with_logits(logits, labels, name=None))
    with tf.variable_scope("hidden", reuse=True):
      loss += 0.5 * tf.reduce_sum(tf.get_variable("weights")**2)
    with tf.variable_scope("output", reuse=True):
      loss += 0.5 * tf.reduce_sum(tf.get_variable("weights")**2)

    return loss

  def accuracy(self, logits, labels):
    correct_prediction = tf.equal(tf.argmax(labels, 1), tf.argmax(logits, 1))
    accuracy = tf.reduce_mean(tf.cast(correct_prediction, tf.float32))
    return accuracy
\end{verbatim}
\end{small}
I also implemented the function \texttt{train} in the file \texttt{train\_mlp.py}:
\begin{small}
\begin{verbatim}
def train():
  # placeholder for input and output tensors allowing to create operations and the computation graph
  x = tf.placeholder(tf.float32, shape=[None, 3072])
  y_ = tf.placeholder(tf.float32, shape=[None, 10])

  # model, namescopes for summary
  model = MLP()
  x_ = model.inference(x)
  with tf.name_scope('loss'):
    loss = model.loss(x_, y_)
  with tf.name_scope('accuracy'):
    accuracy = model.accuracy(x_, y_)
  with tf.name_scope('train_step'):
    train_step = tf.train.GradientDescentOptimizer(LEARNING_RATE_DEFAULT).minimize(loss)

  # session
  init = tf.initialize_all_variables()
  sess = tf.Session()
  sess.run(init)

  # merged summary for the graph, the loss and accuracy
  tf.scalar_summary('accuracy', accuracy)
  tf.scalar_summary('loss', loss)
  tf.histogram_summary("logits", x_) # logits histogram
  merged = tf.merge_all_summaries()

  train_writer = tf.train.SummaryWriter(LOG_DIR_DEFAULT + '/train', sess.graph)
  test_writer = tf.train.SummaryWriter(LOG_DIR_DEFAULT + '/test', sess.graph)

  for i in range(1, MAX_STEPS_DEFAULT+1):
      batch_xs, batch_ys = cifar10.train.next_batch(BATCH_SIZE_DEFAULT)
      __, Summary, l, acc = sess.run([train_step, merged, loss, accuracy],
      feed_dict={x: batch_xs, y_: batch_ys})
      train_writer.add_summary(Summary, i)

      if i % 100 == 0.0:
          batch_xs, batch_ys = cifar10.test.images, cifar10.test.labels
          __, Summary, l, acc = sess.run([train_step, merged, loss, accuracy],
           feed_dict={x: batch_xs, y_: batch_ys})
          test_writer.add_summary(Summary, i)

  train_writer.close()
\end{verbatim}
\end{small}

Figure~\ref{fig:acclosstrain} presents the accuracy and the loss in the training set recorded for every epoch.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{loss.png}\	
\includegraphics[width=0.45\textwidth]{accuracy.png}
\caption{Recorded accuracy and loss for the training set.}
\label{fig:acclosstrain}
\end{figure}

Figure~\ref{fig:acclosstest} presents the accuracy and the loss in the test set recorded for every 100-epochs.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{accuracy_t.png}\	
\includegraphics[width=0.45\textwidth]{loss_t.png}
\caption{Recorded accuracy and loss for the test set.}
\label{fig:acclosstest}
\end{figure}

Figure~\ref{fig:graph} presents the graph of the created model as this is illustrated by TensorBoard.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{graph.png}
\caption{Graph of the model.}
\label{fig:graph}
\end{figure}

Figure~\ref{fig:histogramsunits} presents the histograms of the weights and the biases in every unit of the model, in the hidden and the output units respectively.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{hidden_weights.png}
\includegraphics[width=0.3\textwidth]{hidden_bias.png}
\includegraphics[width=0.3\textwidth]{output_weights.png}
\includegraphics[width=0.3\textwidth]{output_bias.png}
\caption{Histograms of weights and bias in the hidden layer, and weights and bias in the output linear layer.}
\label{fig:histogramsunits}
\end{figure}

Figure~\ref{fig:histogramlogits} presents the histogram of the logits.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{logits.png}
\caption{Histograms of logits.}
\label{fig:histogramlogits}
\end{figure}

\section{Conclusion}
Should contain conclusion of this study.
For example, you can try to answer the following questions.
What was done during this assignment? What features of TensorBoard were positive and what were negative for implementing MLP model and performing the experiments?
What are the main insights you got from the study of the MLP model on CIFAR10 dataset?

\end{document}