### 2016-12-16 practical_3/report/report_methenitis.tex
final submission

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage[final]{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Report for the Deep Learning Course Assignment 3}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Georgios Methenitis \\
  \texttt{georgios.methenitis@cwi.nl}
}

\begin{document}

\maketitle

\begin{abstract}
In this assignment I implemented a convolutional neural network (CNN) with two different architectures on the CIFAR10 dataset.
The first was a straightforward architecture where the input image was fed into a single neural network, while the goal of the neural network was to identify the class of the image.
In the second architecture, two images were the input of two copies of the same neural network, while the goal of the learning was to distinguish between two different images' classes.
In all experiments I used a desktop computer equipped with an NVIDIA GTX770 GPU with 2GB of ram.

\end{abstract}

\section{Task 1}
In this section I explain the implementation and present the results of the first task of the assignment.
I implemented the architecture for the CNN as this was given in the assignment.
For the initialization and regularization of the weights I used:
\begin{verbatim}
initializer = tf.random_normal_initializer( mean=0.0, stddev=0.001 )
regularizer = regularizers.l2_regularizer( 0.001 )
\end{verbatim}
but also experimented with xavier initializer without however reporting the results here, since there was no improvement over the normal distributed initialization of the weights.
Furthermore all biases were initialized at zero:
\begin{verbatim}
initializer=tf.constant_initializer(0.0))
\end{verbatim}
Figure~\ref{fig:1} illustrates the accuracy and the loss during training and test without regularization.
\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-plain.png}\
\includegraphics[height=4.0cm]{loss-plain.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $128$ and no regularization. Green lines illustrate performance on the training set and blue lines on the test set.}
\label{fig:1}
\end{figure}
As shown in the Figures~\ref{fig:1}, the model overfits the training data.
The accuracy in the training set approximates $1.0$ in the last $5000$ steps of the training where the loss in the test set is increasing while accuracy in the test set is not affected.

\subsection{Regularization}
\label{sec:reg}

In this section I present the results of the CNN with regularization losses in the weights.
More specifically I added regularization losses in the fully connected layers $fc1, fc2$
\begin{verbatim}
if self.wd is not None:
    weight_decay = tf.mul(tf.nn.l2_loss(w_fc1), self.wd, name='weight_loss')
    tf.add_to_collection('losses', weight_decay)
\end{verbatim}
making sure the losses were added to the loss function in the loss function of convNet class.
Figure~\ref{fig:2} illustrates the accuracy and the loss during training and test with regularization $wd=0.005$.
\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-plain-reg.png}\
\includegraphics[height=4.0cm]{loss-plain-reg.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $128$ and no regularization. Cyan lines illustrate performance on the training set and orange lines on the test set.}
\label{fig:2}
\end{figure}
There is no performance gain by the regularization apart from a decrease in the loss in the test set.

\subsection{Dropout}

In the next experiment I implemented dropout in the fully connected layers $fc1, fc2$.
\begin{verbatim}
h_fc1 = tf.cond(tf.cast(self.isTrain, tf.bool), 
lambda: tf.nn.dropout(h_fc1, 0.8), lambda: h_fc1)
\end{verbatim}
Figure~\ref{fig:3} illustrates the accuracy and the loss during training and test with dropout $0.5, 0.2$.
\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-plain-drop.png}\
\includegraphics[height=4.0cm]{loss-plain-drop.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $128$ and dropout. Red lines illustrate performance on the training set and purple lines on the test set for dropout $0.2$, while light green lines illustrate performance on the training set and yellow lines on the test set for dropout $0.5$.}
\label{fig:3}
\end{figure}
The performance on the test set in the end of the training was:
\begin{itemize}
\item For no dropout: $\mathbf{0.7503}$
\item For dropout $0.5$: $0.7324$
\item For dropout $0.2$: $0.7433$
\end{itemize}
There was no performance gain, but only a decrease in overfitting with dropout $0.5$.

\subsection{Smaller Batch Size}

Here, I used half the size of the batch to complete training using batch size of $64$.
Figure~\ref{fig:4} illustrates the accuracy and the loss during training and test with the given batch size of $128$ and the half batch size of $64$.

\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-linear-64.png}\
\includegraphics[height=4.0cm]{loss-linear-64.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $64$. Cyan lines illustrate performance on the training set and orange lines on the test set for dropout $0.2$. Green and blue lines illustrate the performance of batch size $128$ with default configuration CNN as in all previous figures.}
\label{fig:4}
\end{figure}
The performance on the test set in the end of the training was:
\begin{itemize}
\item For batch size $128$: $0.7503$
\item For batch size $64$: $\mathbf{0.7520}$
\end{itemize}
There was a performance gain by using smaller batch size and a limited overfitting.

\subsubsection{Regularization}
As in Section~\ref{sec:reg} I used regularization for the weights in the fully connected layers $fc1, fc2$.
Figure~\ref{fig:5} illustrates the accuracy and the loss during training and test with the given batch size of $64$ two regularization values for the weight decays.

\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-linear-64-reg.png}\
\includegraphics[height=4.0cm]{loss-linear-64-reg.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $64$. Cyan lines illustrate performance on the training set and orange lines on the test set without weight regularization. Red lines illustrate performance on the training set and purple lines on the test set with weight regularization $0.01$. Light purple lines illustrate performance on the training set and cyan lines on the test set with weight regularization $0.005$.}
\label{fig:5}
\end{figure}
The performance on the test set in the end of the training was:
\begin{itemize}
\item No regularization: $0.7520$
\item For regularization $0.01$: $\mathbf{0.7634}$
\item For regularization $0.005$: $0.7569$
\end{itemize}
The maximum reported accuracy on the test set was achieved by using mini batches of $64$ size and regularization values of $0.01$.

\subsubsection{Dropout}
Figure~\ref{fig:6} illustrates the results with dropout enables for values $0.2, 0.5$.

\begin{figure}[h!]
\centering
\includegraphics[height=4.0cm]{acc-linear-64-drop.png}\
\includegraphics[height=4.0cm]{loss-linear-64-drop.png}
\caption{Accuracy (left) and loss (right) for the CNN with batch size of $64$. Cyan lines illustrate performance on the training set and orange lines on the test set without weight regularization. Red lines illustrate performance on the training set and purple lines on the test set with weight regularization $0.01$. Light purple lines illustrate performance on the training set and cyan lines on the test set with weight regularization $0.005$.}
\label{fig:6}
\end{figure}

The performance on the test set in the end of the training was:
\begin{itemize}
\item No dropout: $0.7520$
\item For dropout $0.5$: $0.7279$
\item For dropout $0.2$: $0.7493$
\end{itemize}
The was no performance gain by using dropout when using smaller batches of $64$ size.

\subsection{Feature Visualization}
\label{sec:vis}

In this section I show how the features learned during training of the CNN can be illustrated and how they can be used to train an one-vs-all classifier.
I load the learned weights during training and converting them to $2$-d points using TSNE.
\begin{verbatim}
tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
tsne.fit_transform(features)
\end{verbatim}
Figure~\ref{fig:7} presents the visualization of the features in the two dimensional space. 
You can observe that features obtained in $fc2$ layer can distinguish better features of different classes.
On the contrary when see the flatten features all labels are located close to points of other classes.
\begin{figure}[h!]
\centering
\includegraphics[height=5.0cm]{visualization-linear-features_flatten.pdf}
\includegraphics[height=5.0cm]{visualization-linear-features_fc1.pdf}
\includegraphics[height=5.0cm]{visualization-linear-features_fc2.pdf}
\caption{Visualization of the features in the 2-d space of the \texttt{flatten} layer (top), \texttt{fc1} (bottom-left), \texttt{fc2} (bottom-right).}
\label{fig:7}
\end{figure}

\subsubsection{One-vs-Rest Classifier}

To test the accuracy of the one-vs-all classifier I use the features retained by the learned model to train the classifiers.
Using the 80 percent of the features I train a linear kernel OneVsRest classifier, and test the accuracy over the rest 20 percent.
\begin{verbatim}
train_set = range(int(features.shape[0]*0.8))
test_set = range(train_set[-1] + 1, features.shape[0])

classif = OneVsRestClassifier(SVC(kernel='linear'))
classif.fit(features[train_set], batch_y[train_set]
classif.score(features[test_set], batch_y[test_set]
\end{verbatim}
Here, the reported accuracy of the one-vs-classifiers on the test set:
\begin{itemize}
\item Accuracy with flatten features: $0.6305$.
\item Accuracy with fc1 features: $0.7465$.
\item Accuracy with fc2 features: $\mathbf{0.7685}$.
\end{itemize}
It is natural that the accuracy is higher for later layers.
As it was earlier shown $fc2$ layer can generates features that the one-cs-rest classifier can distinguish different classes easier than from the other layers.

\section{Task 2}

In this section I present my implementation and the results achieved by the siamese architecture of the CNN.		
First, I implemented the functions \texttt{create\_dataset} and \texttt{next\_batch} in the \texttt{cifar10\_siamese\_utils.py}.

\subsection{Dataset}

The create\_dataset function:
\begin{tiny}
\begin{verbatim}
dset = []
    for sample_tuple in range(num_tuples):

        n_correct = int(fraction_same * batch_size)

        val_set_indexes = range(int(source_data.train._images.shape[0] * 0.8), source_data.train._images.shape[0])

        index_x1 = np.random.choice(val_set_indexes)
        label_x1 = source_data.train._labels[index_x1]

        index_x2_similar = np.random.choice(val_set_indexes[0] +
                                            np.where(source_data.train._labels[val_set_indexes] == label_x1)[0],
                                            replace=False, size=n_correct)

        index_x2_opposite = np.random.choice(val_set_indexes[0] +
                                             np.where(source_data.train._labels[val_set_indexes] != label_x1)[0],
                                             replace=False, size=(batch_size - n_correct))

        x1 = np.array([source_data.train._images[index_x1] for i in range(batch_size)])
        x2 = source_data.train._images[index_x2_similar]
        x2 = np.vstack((x2, source_data.train._images[index_x2_opposite]))
        labels = np.hstack((np.ones((n_correct)), np.zeros((batch_size - n_correct))))

        dset.append((x1, x2, labels))

    return dset
\end{verbatim}
\end{tiny}
which returns \texttt{num\_tuples} tuples of the following batches:
\begin{verbatim}
      X_1            X_2               Y
| image_cl1_1, image_cl1_10  | -->   | 1 |
| image_cl1_1, image_cl1_4   | -->   | 1 |
| image_cl1_1, image_cl1_163 | -->   | 1 |
| image_cl1_1, image_cl1_145 | -->   | 1 |
| image_cl1_1, image_cl3_8   | -->   | 0 |
|      .            .        | -->   | 0 |
|      .            .        | -->   | 0 |
|      .            .        | -->   | 0 |
| image_cl1_1, image_cl5_8   | -->   | 0 |
| image_cl1_1, image_cl2_    | -->   | 0 |
| image_cl1_1, image_cl10_8  | -->   | 0 |
\end{verbatim}
and used for the evaluation of the training set.

The \texttt{next\_batch} function, which include the following piece of code was given to us in \texttt{cifar10\_utils.py} to sample as uniformly the dataset as possible.
\begin{tiny}
\begin{verbatim}
n_correct = int(fraction_same * batch_size)
        val_set_indexes = range(int(self._images.shape[0] * 0.8))

        start = self._index_in_epoch
        self._index_in_epoch += batch_size
        if self._index_in_epoch > self._num_examples * 0.8:
            self._epochs_completed += 1
            np.random.shuffle(val_set_indexes)
            start = 0
            self._index_in_epoch = batch_size
            assert batch_size <= self._num_examples

        end = self._index_in_epoch

        val_set_indexes = val_set_indexes[start:end]
\end{tiny}
\end{verbatim}
\end{tiny}
The \texttt{next\_batch} function then creates each training batch in two following ways.
\begin{itemize}
\item In the \emph{default} implementation the structure of each batch is the same as the test set following the same idea.
All images in $x1$ are the same, while $x2$ contains \texttt{fraction} percentage of similar label images to $x1$ and the rest random images of different classes.
\item In the \emph{alternative} implementation the following piece of code
\begin{tiny}
\begin{verbatim}
random_index = np.array(np.random.choice(val_set_indexes, replace=False, size=batch_size))

matches = [np.random.choice(np.array(np.where(self._labels[range(int(self._images.shape[0] * 0.8))] == 
self._labels[x]))[0]) for x in random_index]

mismatches = [np.random.choice(np.array(np.where(self._labels[range(int(self._images.shape[0] * 0.8))] != 
self._labels[x]))[0]) for x in random_index]

x1 = self._images[random_index]
x2 = self._images[matches[:n_correct]]
x2 = np.vstack((x2, self._images[mismatches[n_correct:]]))
labels = np.hstack((np.ones((n_correct)), np.zeros((batch_size-n_correct))))
\end{verbatim}
\end{tiny}
generates batches of random pairs of images respecting the \texttt{fraction} of same label images.
The resulted batch 
\begin{verbatim}
      X_1            X_2               Y
| image_cl1_10, image_cl1_50   | -->   | 1 |
| image_cl4_17, image_cl4_234  | -->   | 1 |
| image_cl8_654, image_cl8_14  | -->   | 1 |
| image_cl8_66, image_cl6_234  | -->   | 0 |
|      .            .          | -->   | 0 |
|      .            .          | -->   | 0 |
|      .            .          | -->   | 0 |
| image_cl2_327, image_cl6_234 | -->   | 0 |
\end{verbatim}
Is is expected that classes are uniformly picked in expectation given the random sampling. Given the default way of selecting training batches training among classes is not guarranteed to be uniform.
Later we see that the alternative way of creating batches achieves the highest accuracy in the end of training using the one-vs-rest classifier.
\end{itemize}

\subsection{Loss function}

I used the following two loss functions:
\begin{enumerate}
\item $\mathcal{L} =  Y * \frac{1}{2} * d^2 + (1-Y) * max(margin - d^2, 0)$
\item $\mathcal{L} =  Y * d^2 + (1-Y) * max(margin - d^2, 0)$
\end{enumerate}
In the first the distance in similar label images is multiplied by half, while in the second is not.

\subsection{Training}

I construct 3 experiments using both the \texttt{next\_batch} ways (default \& alternative) and the two loss functions.
\begin{description}
\item[Experiment 1]{Using the loss function (1) and the default creation of batches}
\item[Experiment 2]{Using the loss function (2) and the default creation of batches}
\item[Experiment 3]{Using the loss function (2) and the alternative creation of batches}
\end{description}
Figure~\ref{fig:siamese} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 1.
Loss in the test set is computed every $1000$ steps.
We see that the loss in the training set in decreasing as the training steps increase.
The same occurs for the loss in the test set.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese.png}
\caption{\textbf{Experiment 1} Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese}
\end{figure}
Figure~\ref{fig:siamese-1} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 2.
Here the scaling of the loss is different as a result of the different loss function used in experiment 2.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-1.png}
\caption{\textbf{Experiment 2} (Different Loss Functions) Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese-1}
\end{figure}
Figure~\ref{fig:siamese-2} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 3.
Here I compare between two different ways of batch sampling between experiment 2 and 3.
The same loss function in used both experiments.
It is easy to see that using the alternative batch sampling, the loss is decreasing faster and results in lower loss in the test set as well.
In the next section I show that the features retained with the alternative batch sampling increase the accuracy of the one-vs-rest classifier.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-2.png}
\caption{\textbf{Experiment 3} (Default vs Alternative Batch Creation) Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (yellow line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).}
\label{fig:siamese-2}
\end{figure}

\subsection{Feature Visualization}

In the previous section I set up three experiments and presented the loss during training in the training set.
In this section I illustrate the features in the 2-dimensional space as these transformed by TSNE (as this was implemented in Section~\ref{sec:vis}).
Figure~\ref{fig:features} presents the features generated by the normalized output layer of the network for the setting used in the 3 experiments.

\begin{figure}[h!]
\centering
\includegraphics[width=6.2cm]{visualization-siamese-plain.pdf}
\includegraphics[width=6.2cm]{visualization-siamese-1.pdf}
\includegraphics[width=7.2cm]{visualization-siamese-3.pdf}
\caption{Visualization of the $2D$ features generated out of the whole test set. Experiment 1 (top left), experiment 2 (top right), experiment 3 (bottom).}
\label{fig:features}
\end{figure}

\subsection{One-vs-Rest Classifier}

Here, I present the accuracy of the one-vs-rest classifier on the features learned from in the final layer of the siamese network for all experiments.

\begin{itemize}
\item \textbf{Experiment 1}: $0.476$
\item \textbf{Experiment 2}: $0.473$
\item \textbf{Experiment 3}: $\mathbf{0.682}$
\end{itemize}

Using the different loss function with the default batch sampling method did not help the accuracy obtained.
However, in experiment 3 I used the alternative batch sampling function which was the main reason for the large improvement in the accuracy.

\clearpage

\section{Task 3: Transfer Learning}

\begin{figure}[h!]
\centering
\includegraphics[width=5.2cm]{graph.png}
\caption{Graph for the task 3 of the assignment, the final three layers of convnet were used. Note the stop gradient operation, where we don't want to update the pre-trained weights.}
\label{fig:graph}
\end{figure}

\subsection{Feature Extraction}

In this section I present the results of my implementation for the feature extraction of the third task.
I used the last 3 layers of the convNet inference method from the first task of the assignment.
I also used stopGradient method of tensorflow to disconnect vgg pretrained network from convNet and the training updates.
Figure~\ref{fig:extra} illustrates the performance on the training and the test set for $10000$ training steps.
The accuracy, when updating only the last three layers of the network, approximates $0.67$ which is lower than the performance of the network in task 1.
It shows that features obtained by another pretrained network cannot work that well in an another task.

\begin{figure}[h!]
\centering
\includegraphics[width=12.2cm]{acc-transfer.png}\	
\includegraphics[width=12.2cm]{loss-transfer.png}
\caption{(Feature extraction) Accuracy and loss in the training (purple line) and the test set (yellow line) for $10000$ training step.}
\label{fig:extra}
\end{figure}

\subsection{Retraining}

Here I present the results of retraining, stop gradient is disabled and thus all the weights are updated from the beginning of the training procedure.
Figure~\ref{fig:retrain} illustrates the performance on the training and the test set for $10000$ training steps.
The accuracy approximates $0.85$ which is $9\%$ higher than the best performance achieved in task 1 of the assignment.
Which is a natural result given that weights are initialized with ``good'' priors.

\begin{figure}[h!]
\centering
\includegraphics[width=12.2cm]{acc-retrain.png}\	
\includegraphics[width=12.2cm]{loss-retrain.png}
\caption{(Retraining) Accuracy and loss in the training (orange line) and the test set (brown line) for $10000$ training step.}
\label{fig:retrain}
\end{figure}

\subsubsection{Refining}

Here I present the results of refining, stop gradient is disabled after $k$-training steps and thus all the weights are updated only after $k$-step during training.
Figure~\ref{fig:refine} illustrates the performance on the training and the test set for $10000$ training steps.
The accuracy approximates $0.85$ which is equivalent to the performance achieved in the retraining task.
Note, the big drops in accuracy and jumps in loss after the refining starting step.
This is due to the heavy influence the initial layers of the VGG have on the classification, where small changes in the weights influence a lot the output of the network.

\begin{figure}[h!]
\centering
\includegraphics[width=12.2cm]{acc-refine.png}\	
\includegraphics[width=12.2cm]{loss-refine.png}
\caption{(Refining) Accuracy and loss in the training (orange line) and the test set (blue line) for $10000$ training step, refining after $100$ training steps.
Accuracy and loss in the training (purple line) and the test set (cyan line) for $10000$ training step, refining after $1000$ training steps.
Accuracy and loss in the training (green line) and the test set (blue line) for $10000$ training step, refining after $2500$ training steps.}
\label{fig:refine}
\end{figure}

\subsection{Performance of transfer learning on Cifar10}

Here I present the results on the test set after $10000$ training steps of all the methods studied above.
\begin{description}
\item[Feature extraction:]{$0.6743$}
\item[Retraining:]{$0.8505$}
\item[Refining-$100$:]{$0.8441$}
\item[Refining-$1000$:]{$0.8461$}
\item[Refining-$2500$:]{$\mathbf{0.8510}$}
\end{description}
The highest performance was achieved by refining-$2500$ where the weights of the pre-trained network started to be optimized for the Cifar10 task after $2500$ training steps.
However, the performance was approximately equivalent to the retraining method, where all weights are updated during the whole period of training.

\section{Conclusion}

In this assignment I implemented a CNN and reported its performance over the CIFAR10 dataset using different parameters.
I also implemented the siamese architecture which uses two versions of the same CNN to perform learning in order to distinguish between classes.
I showed that using the alternative batch sampling procedure the loss is reduced and the features retained result in better accuracy when an one-vs-rest classifier is used.
The amount of training steps requires for the siamese are higher than the single CNN architecture.
Last, I implemented the third task of the assignment where we import pre-trained layers from VGG-16 and apply learning on Cifar10 dataset.

\end{document}